---
title: "Processing Depth Data"
author: "Kourtney Burger"
date: "2024-01-18"
output: html_document
---
```{r}
library(tidyverse)
library(chron)
library(lubridate)
```

# OVERVIEW
1. Read in raw depth data
2. Add column names and additional depth/time columns
3. Fill in columns to calculate depth and date/
4. Remove unnecessary columns and export final data

# ADRIFT Sensus Depth Data
## 1. Read in raw depth data
```{r}
#set temporary working directory
setwd('Z:\\METADATA\\ADRIFT\\ADRIFT_108\\ADRIFT_108_Depth_Sensus')
#Read in data
#rawDepth <- read.csv('ADRIFT_108_Bottom_Depth_Sensus.csv', header = FALSE)
rawDepth <- read.csv('ADRIFT_108_Top_Depth_Sensus.csv', header = FALSE)
#rawDepth <- read.csv('U-16656.csv', header = FALSE)
#write.csv(rawDepth, "Z:\\METADATA\\ADRIFT\\ADRIFT_108\\ADRIFT_108_Depth_Sensus\\ADRIFT_108_Bottom_Depth_Sensus_RAW.csv", row.names = FALSE)
write.csv(rawDepth, "Z:\\METADATA\\ADRIFT\\ADRIFT_108\\ADRIFT_108_Depth_Sensus\\ADRIFT_108_Top_Depth_Sensus_RAW.csv", row.names = FALSE)
```

## 2. Add column names and additional depth/time columns
```{r}
#Add other columns
rawDepth <- rawDepth %>%
  add_column(Depth_m = NA, Temp_C = NA, Start_Time_UTC = NA, Start_Date_UTC = NA, DateTime_UTC = NA)
#Add column names
colnames(rawDepth) <- c('Leg','Sensus ID','','Year','Month','Day','StartHr UTC','StartMin UTC','StartSec UTC','Elapsed Sec','mBar','degK','Depth_m','Temp_C','Start_Time_UTC','Start_Date_UTC','DateTime_UTC')
```

## 3. Fill in columns to calculate depth and date/time
```{r}
# Calculate depth using pressure (mBar)
for (i in 1:nrow(rawDepth)) {
 rawDepth$Depth_m[i] <- (rawDepth$mBar[i]-992)/103
}

# Convert temperature from Kelvin to Celsius
for (i in 1:nrow(rawDepth)) {
  rawDepth$Temp_C[i] <- rawDepth$degK[i]-273.15
}

# Add start date column
for (i in 1:nrow(rawDepth)) {
  rawDepth$Start_Date_UTC[i] <- paste(rawDepth$Year[i],"-",rawDepth$Month[i],"-",rawDepth$Day[i], sep = "")
}

# rawDepth$Year <- as.integer(rawDepth$Year)
# rawDepth$Month <- as.integer(rawDepth$Month)
# rawDepth$Day <- as.integer(rawDepth$Day)
#
# rawDepth %>%
#   select(Year, Month, Day) %>%
#   mutate(Start_Date_UTC = make_date(Year, Month, Day))

# Add start time column
for (i in 1:nrow(rawDepth)) {
  rawDepth$Start_Time_UTC[i] <- paste(rawDepth$`StartHr UTC`[i],":",rawDepth$`StartMin UTC`[i],":",rawDepth$`StartSec UTC`[i], sep = "")
}

# Add date/time column
for(i in 1:nrow(rawDepth)) {
  rawDepth$DateTime_UTC[i] <- paste(rawDepth$Start_Date_UTC[i]," ",rawDepth$Start_Time_UTC[i])
}

rawDepth$DateTime_UTC <- as.POSIXct(rawDepth$DateTime_UTC, format = "%Y-%m-%d %H:%M:%S")

for (i in 1:nrow(rawDepth)) { 
  rawDepth$DateTime_UTC[i] <- rawDepth$DateTime_UTC[1] + rawDepth$`Elapsed Sec`[i]
}
```

## 4. Remove unnecessary columns and export final data
```{r}
finalDepth <- subset(rawDepth, select = c("Sensus ID","DateTime_UTC","mBar","degK","Depth_m","Temp_C"))
```

```{r}
#write.csv(finalDepth, "Z:\\METADATA\\ADRIFT\\ADRIFT_108\\ADRIFT_108_Depth_Sensus\\ADRIFT_108_Bottom_Depth_Sensus.csv", row.names=FALSE)
write.csv(finalDepth, "Z:\\METADATA\\ADRIFT\\ADRIFT_108\\ADRIFT_108_Depth_Sensus\\ADRIFT_108_Top_Depth_Sensus.csv", row.names=FALSE)
```


# CCES/PASCAL OpenTag Depth Data

## 1. Read in raw depth data
```{r}
#set temporary working directory
setwd('Z:/METADATA/PASCAL/PASCAL_002')

#Rename raw data
file.rename('PASCAL_002_OT_Depth.csv', 'PASCAL_002_OT_Depth_Raw.csv')

#Read in raw data
rawDepth <- read.csv('PASCAL_002_OT_Depth_Raw.csv', header = FALSE)
```

## 2. Add column names and additional depth/time columns
```{r}
#Add other columns
rawDepth <- rawDepth %>%
  add_column(Depth_m = NA, DateTime_UTC = NA)

#Add column names
colnames(rawDepth) <- c('Num','FileName','FileDate','FileTime','Sample','TimeFromStart_s','Pressure','Temperature_c','Depth_m', 'DateTime_UTC')
```

## 3. Fill in columns to calculate depth and date

```{r}
# Set up progress bar
pb <- txtProgressBar(min = 0,      # Minimum value of the progress bar
                     max = nrow(rawDepth), # Maximum value of the progress bar
                     style = 3,    # Progress bar style (also available style = 1 and style = 2)
                     width = 50,   # Progress bar width. Defaults to getOption("width")
                     char = "=")   # Character used to create the bar

# Calculate depth using pressure (mBar)
for (i in 1:nrow(rawDepth)) {
  rawDepth$Depth_m[i] <- (rawDepth$Pressure[i] - 1016)/100.5 
  
  # Sets the progress bar to the current state
  setTxtProgressBar(pb, i)
}

close(pb) # Close the connection

# Convert date and time into datetime object
  dates <- rawDepth$FileDate
  times <- rawDepth$FileTime
  x <- paste(dates, times)
  
  rawDepth$DateTime_UTC <- as.POSIXct(x, format = "%m/%d/%Y %H:%M:%S")
  
# # Correct time zone as needed (-7 hours)
# for (i in 1:nrow(rawDepth)) {
#   rawDepth$DateTime_UTC[i] <- rawDepth$DateTime_UTC[i]-(7*3600)
# 
#   # Sets the progress bar to the current state
#   setTxtProgressBar(pb, i)  
# }
# 
# close(pb) # Close the connection  
```

## 4. Remove unnecessary columns and export final data
```{r}
finalDepth <- subset(rawDepth, select = c("FileName","DateTime_UTC","Pressure","Depth_m","Temperature"))
```

```{r}
write.csv(finalDepth, "Z:/METADATA/PASCAL/PASCAL_002/PASCAL_002_Depth_OpenTag/PASCAL_002_OT_Depth.csv", row.names=FALSE)
```


# CCES LOTEK Depth Data

## 1. Read in raw depth data
```{r}
#set temporary working directory
setwd('Z:/METADATA/CCES_2018/CCES_014/CCES_014_DEPTH/')

#Rename raw data
file.rename('CCES_014_LAT_Depth.csv', 'CCES_014_LAT_Depth_Raw.csv')

#Read in raw data
rawDepth <- read.csv('CCES_014_LAT_Depth_Raw.csv', header = FALSE)

# NOTE: IF WORKING WITH LOTEK DATA YOU MUST REMOVE THE FIRST THREE ROWS OF THE RAW DATA
rawDepth <- rawDepth[-c(1,2,3), ]
```

## 2. Add column names and additional depth/time columns
```{r}
#Add other columns
rawDepth <- rawDepth %>%
  add_column(Depth_m = NA, DateTime_UTC = NA)

#Add column names
colnames(rawDepth) <- c('Rec','Date','Time','Pressure','IntTemp','Depth_m', 'DateTime_UTC')
```

## 3. Fill in columns to calculate depth and date

```{r}
# Set up progress bar
pb <- txtProgressBar(min = 0,      # Minimum value of the progress bar
                     max = nrow(rawDepth), # Maximum value of the progress bar
                     style = 3,    # Progress bar style (also available style = 1 and style = 2)
                     width = 50,   # Progress bar width. Defaults to getOption("width")
                     char = "=")   # Character used to create the bar

# Calculate depth using pressure (mBar)

# NOTE: IF USING LOTEK DATA YOU MUST CONVERT PRESSURE FROM dBar to mBar
rawDepth$Pressure <- as.numeric(rawDepth$Pressure)*100

for (i in 1:nrow(rawDepth)) {
  rawDepth$Depth_m[i] <- (rawDepth$Pressure[i] - 1016)/100.5 
  
  # Sets the progress bar to the current state
  setTxtProgressBar(pb, i)
}

close(pb) # Close the connection

# Convert date and time into datetime object
  dates <- rawDepth$Date
  times <- rawDepth$Time
  x <- paste(dates, times)
  
  rawDepth$DateTime_UTC <- as.POSIXct(x, format = "%m/%d/%Y %H:%M:%S")
  
# # Correct time zone as needed (-7 hours)
# for (i in 1:nrow(rawDepth)) {
#   rawDepth$DateTime_UTC[i] <- rawDepth$DateTime_UTC[i]-(7*3600)
# 
#   # Sets the progress bar to the current state
#   setTxtProgressBar(pb, i)  
# }
# 
# close(pb) # Close the connection  
```

## 4. Remove unnecessary columns and export final data
```{r}
finalDepth <- subset(rawDepth, select = c("Rec","DateTime_UTC","Pressure","Depth_m","IntTemp"))
```

```{r}
write.csv(finalDepth, "Z:/METADATA/CCES_2018/CCES_014/CCES_014_DEPTH/CCES_014_LAT_Depth.csv", row.names=FALSE)
```
